---
title: "Estimating small mammal and reptile population size using Mark-Recapture"
subtitle: "ESM 212 Assignment 3"
author: "Your name here"
date: "2025-03-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r packages}
library("ggplot2")
library("tidyverse")
install.packages("PASWR")
library("PASWR")
install.packages("Rcapture")
library(Rcapture)

```



The purpose of this exercise is to reinforce the course field trip and readings estimating animal population size using the closed-population mark-recapture method. Please read the instructions carefully. You can knit this to a pdf and submit it to canvas, or insert your answers into the word document and then upload to canvas. Please note: this is meant to be useful and interesting, not to simply make work for you. Please let me know if you have any questions or if the instructions are not clear and I will try to clarify asap.

#### Methods

Methods
Over the course of this week, we simulated a mark-recapture study of small mammals at Coal Oil Point, sampling for two consecutive nights on a regular trapping grid. Individuals would have been marked with ear tags to be able to identify recaptures. For the sake of this exercise, we will use a subset of the data from MacDonald et al. 2018 Ecosphere, as the class case study data are not sufficient to make population estimates. We will assume that the populations were closed: no births, deaths, immigration, or emigration, which is reasonable given the short timeframe of the sampling and the localized area in which sampling took place. In this case, three consecutive trapping nights were used to distinguish population sizes between burned and unburned plots in Los Padres National Forest following the 2013 White Fire. In this analysis we will use a Huggins closed population model, making the simplest assumptions about capture and recapture probabilities (e.g. they are equal and do not differ based on age, sex, prior capture history, etc.)

#### Data 

The data can be found in csv files in the exercise folder (FYI, no N. fuscipes were trapped in the burned plots, so there is no data file for this):

```{r data}

# Unburned peromyscus data
PM.unburned <- read.csv("data/PM_unburned.csv", header=T)
PM.unburned$X <- NULL
PM.unburned <- data.frame(PM.unburned)
PM.unburned

# Burned peromyscus data
PM.burned <- read.csv("data/PM_burned.csv")
PM.burned$X <- NULL
PM.burned <- data.frame(PM.burned)
PM.burned


```


## 1a) What is the estimated population size of Peromyscus mice inside and outside the burn area (this is done in the code cell below for you)? (Use a simple closed population function from the R package `Rcapture`).
```{r}

# Fit closed population models using Rcapture
PMPop.unburned <- closedp(PM.unburned)


# Extract the estimate from the M0 model
N_hat_M0 <- PMPop.unburned$results["M0", "abundance"]


# Display the results for both models
cat("Population estimate:", N_hat_M0)



```

## 1b) Repeat 1a for Neotoma fuscipes (woodrats) and Sceloporus occidentalis (fence lizards).
```{r}

```



## 2a) Use a t-test to compare population estimates for Peromyscus mice between the burned and unburned plots (this is done for you in the r script).

```{r}
# compare P. maniculatus populations between unburned (x) and burned (y)
PMt.test <- tsum.test(mean.x=12.8397, s.x=3.956004, n.x=9, mean.y=11.5019, s.y=5.541892, n.y=9, alternative="two.sided", mu=0, var.equal=T, conf.level=0.95)
PMt.test # No, not significantly different n.x=9, n.y=9: 3 grids trapped for 3 nights each
```


## 2b) Repeat 2a for Neotoma fuscipes (woodrats) and Sceloporus occidentalis (fence lizards).
```{r}
#HINT: set mean = 0, SD = 0 and sample size (n) = 9 (for three trapping grids sampled, 3 nights each), for the burned sites where no animals were captured (in order to estimate differences with the t-test approach)

```

## 3) From what you have learned about mark-recapture and animal behavior, briefly discuss the main sources of uncertainty in such a mark-recapture survey and how you might change our method (sampling or modeling approach) to reduce these uncertainties.

*Your answer here.*

